% \input{sections/par-organization.tex}

Pattern recognition is an applied science that draws from a variety of mathematical fields.
Principal component analysis (PCA) is an unsupervised learning technique that can be used to reveal patterns in high-dimensional data.
The transformation used in the PCA algorithm resembles the singular value decomposition (SVD) in linear algebra and the Karhunen-Loève transform (KLT) in stochastic processes.
Borrowing results from these different approaches, we can start to generalize the PCA algorithm.
The kernel trick from machine learning can be used to create a nonlinear algorithm from a linear one.
In particular, the kernel trick can be applied to PCA.
Kernel PCA is a nonlinear version of PCA that maps data to a suitable Hilbert space and attempts to find patterns in this new context.

We will begin this paper by deriving the PCA transform in \Cref{sec:principal-component-analysis}.
Part of this work depends on results from the SVD and is discussed in \Cref{sub:singular-value-decomposition}.
The PCA section is concluded with examples.
In \Cref{sec:reproducing-kernel-hilbert-space}, we justify the kernel trick.
This prompts the discussion of reproducing kernel Hilbert spaces, Mercer's theorem, the Moore-Aronszajn theorem, and the Riesz representation theorem.
Last, the kernel PCA algorithm is derived in \Cref{sec:kernel-pca}.

% \input{sections/par-linear-regression.tex}

% \begin{example}
%     \label{eg:regression}
%     \input{sections/def-thm/eg-regression.tex}
% \end{example}